### 2014-03-20 manuscript.tex
minor typos etc

Social media and online discussions contain a wealth of information which can make us able to understand the extent of hate speech on the Internet. However, it turns out that academia is lacking research on social media and online radicalization \cite{s1}. Opinion mining is the discipline of automatically extracting opinions from a text material and may be one important tool in the understanding online radicalization. Opinion mining has mostly been used to analyze opinions in comments and reviews about commercial products, but there are also examples of opinion mining towards political tweets and discussions, see e.g. Tumasjan et al. \shortcite{Tumasjan2010}; Chen et al. \shortcite{Chen10}. Opinion mining towards political discussions is known to be hard since citations, irony and sarcasm is very common \cite{Bing12}.

Opinion classification is perhaps the most studied topic within opinion mining. It aims to classify a set of text as either positive or negative and sometimes also neutral. There are mainly two approaches, one based on machine learning and one based on based on using a list of words with given sentiment scores (lexical approach). One simple lexical approach is to count the number of words with positive and negative sentiment in the document as suggested by Hu and Liu \shortcite{Hu04}. One may classify the opinion of larger documents like movie or product reviews or smaller documents like tweets, comments or sentences. See Liu \shortcite{Bing12}, chapters three to five and references therein for the description of several opinion classification methods.

@@

In this Section we present two methods to classify sentences to either positive, neutral or negative towards a keyword. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Both keywords, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters is words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. Let $kw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the keyword in the sentence. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ repressent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(kw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function \textbf{imp} computes the importance of the sentiment word $sw_{j}$ on the keyword appearance $kw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters is within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the opinion towards the keyword to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set.

@@

\paragraph{Modified scoring function}
Let $\mathcal{D}$ denote the set of all important grammatical relations. The function $\mathbf{gram}(kw_i, sw_{j})$ returns the grammatical relation, and if $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{j}) \in [0,1]$, returns the importance of the grammatical relation. Further let $\mathbf{depdist}(kw_i, sw_{j})$ return the dependency distance, as described above. The \textbf{imp} function is computed as follows. If $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(kw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(kw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is a parameter that weights the score from the important dependence path and the tree distance and can be tuned using a training set. If $\mathbf{gram}(kw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(kw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.

@@

We did not find any suitable annotated text material related to political discussions and therefore created our own. We manually selected 46 debate articles from the Norwegian online newspapers \textit{NRK Ytring, Dagbladet, Aftenposten, VG and Bergens Tidene}. To each debate article there were attached a discussion thread where readers could express their opinions and feelings towards the content of the debate article. All the text from the debate articles and the subsequent discussions were collected using text scraping \cite{Hammer13}. The debate articles were related to religion and immigration and we wanted to classify the opinion towards all words with stem: \textit{islam, muslim, quran, allah, muhammed, imam and mosque}. These are topics that typically creates a lot of active discussions and disagreements.

@@

Each sentence were manually annotated whether the commenter were positive, negative or neutral towards the keyword in the sentence. Each sentence were evaluated individually. The sentences were annotated based on all our knowledge, e.g. a sentence like ``Muhammed is like Hitler'' would be annotated as a negative opinion towards Muhammed. Further, if a commenter presented a negative fact about the keyword, the sentence would be denoted as negative.

A random sample of 65 sentences from the original text material were annotated by a second annotator. These sentences were not included in either the training or test set. For these sentences, the two annotators agreed on 58 of them, which is an 89\% agreement, with a 95\% confidence interval equal to (79\%, 95\%) assuming that each sentence is independent. Since the sentences is drawn randomly from the population of all sentences this is a fair assumption.

Finally the material were divided in to two parts where the first half of the debate articles with subsequent discussions where in the training set and the rest were in the test set. The researcher working on finding the important dependence paths, did only use the training set and did never see the test set before the decisions about dependence paths were decided. After the division, the training and test set consisted of a total of 382 and 308 sentences, respectively. Table \ref{tab:1} summarizes the opinions in the sentences
\begin{table}
  \caption{Manual annotation of training and test set.}
  \centering
  \begin{tabular}{lcccc}
             & Negative & Neutral & Positive \\\hline
    Training & 174 (46\%) & 162 (42\%) & 46 (12\%)\\
    Test     & 102 (33\%)& 182 (59\%) & 24 (8\%)
  \end{tabular}
  \label{tab:1}
\end{table}

@@

In this study we compared four different methods based on the general algorithm in \eqref{eq:1}.
\begin{itemize}
\item We use the \textbf{imp}-function presented in \eqref{eq:2}. We denote this method WD (word distance).
\item For this method and the two below we use the \textbf{imp}-function in \eqref{eq:3}. Further we set $\alpha = 0$ which means that we do not use the important dependence paths. We denote this method A0 ($\alpha = 0$).
\item We set $\alpha = 1$ and for all dependency paths we set $W_{\mathbf{dep}} = 2/3$. We denote this method CW (constant weights).
\item We set $\alpha = 1$ and for $W_{\mathbf{dep}}$ we use the weights presented in Table ??????. We denote this method OD (optimal use of dependence paths)
\end{itemize}
For each method we used the training set to manually tune the parameters $d_p, d_n, t_p$ and $t_n$ of the method. The parameters were tuned to optimize the number of correct classifications.

Table \ref{tab:2} shows the optimal parameter values of $d_p, d_n, t_p$ and $t_n$ tuned using the training set, and classification performance for the different methods on the test set using the parameter values tuned from the training set. The p-values are computed using the regression model presented in Section \ref{sec:sa}.
\begin{table}
  \caption{The second to the fifth column show the optimal values of the parameters of the model tuned using the training set. The sixth column show the number of correct classifications and the last column shows p-values testing whether the method performs better than WD.}
  \centering
  \begin{tabular}{lcccccl}
       & $d_p$ & $d_n$ & $t_p$ & $t_n$ & Correct   & p-val \\ \hline
    WD &    2  &   0  & 0.7  &  0.0  & 145 (47\%) & \\
    A0 &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.023\\
    CW &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.024\\
    OD &    2  &   0  & 2.0  &  0.3  & 162 (53\%) & 0.016
  \end{tabular}
  \label{tab:2}
\end{table}
We see that the sentiment shifter 'not' only have a positive effect on the classification performance when it is in front of the sentiment word. We see that using dependence tree distances (method A0) the classification results is significantly improved compared to using word distances in the sentence (method WD) (p-value = 0.023). Also classification based on important dependence paths (method OD) performs significantly better than WD. We also see that OD performs better than A0 (162 correct compared to 161), but this improvement is not statistically significant.

@@

Classifying opinions in political discussions is hard because of the frequent use of irony, sarcasm and citations. In this paper we have compared the use of word distance between keyword and sentiment word against metrics related to parsed sentence information. Our results show that using dependence tree distances or important dependence paths, improves the classification performance compared to using word distance.

Manually selecting important dependence paths for the aim of opinion mining is a hard task. A natural further step of our analysis is to expand the training and test material and use machine learning to see if there exists dependence paths that improve results compared to using dependence tree distance.

>>>>

Social media and online discussions contain a wealth of information which can make us able to understand the extent of hate speech on the Internet. However, it turns out that academia is lacking research on social media and online radicalization \cite{s1}. Opinion mining is the discipline of automatically extracting opinions from a text material and may be one important tool in the understanding of online radicalization. Opinion mining has mostly been used to analyze opinions in comments and reviews about commercial products, but there are also examples of opinion mining towards political tweets and discussions, see e.g. Tumasjan et al. \shortcite{Tumasjan2010}; Chen et al. \shortcite{Chen10}. Opinion mining towards political discussions is known to be hard since citations, irony and sarcasm is very common \cite{Bing12}.

Opinion classification is perhaps the most studied topic within opinion mining. It aims to classify a set of text as either positive or negative and sometimes also neutral. There are mainly two approaches, one based on machine learning and one based on based on using a list of words with given sentiment scores (lexicon-based approach). One simple lexicon-based approach is to count the number of words with positive and negative sentiment in the document as suggested by Hu and Liu \shortcite{Hu04}. One may classify the opinion of larger documents like movie or product reviews or smaller documents like tweets, comments or sentences. %% See Liu \shortcite{Bing12}, chapters three to five and references therein for the description of several opinion classification methods.

@@

In this Section we present two methods to classify sentences as either positive, neutral or negative towards a keyword. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Both keywords, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters are words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. Let $kw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the keyword in the sentence. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ repressent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(kw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function \textbf{imp} computes the importance of the sentiment word $sw_{j}$ on the keyword appearance $kw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters is within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the opinion towards the keyword to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set.

@@

\paragraph{Modified scoring function}
Let $\mathcal{D}$ denote the set of all important dependency paths. The function $\mathbf{gram}(kw_i, sw_{j})$ returns the dependency path, and if $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{j}) \in [0,1]$, returns the importance of the path. Further let $\mathbf{depdist}(kw_i, sw_{j})$ return the dependency distance, as described above. The \textbf{imp} function is computed as follows. If $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(kw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(kw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is a parameter that weights the score from the important dependency path and the tree distance and can be tuned using a training set. If $\mathbf{gram}(kw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(kw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.

@@

We did not find any suitable annotated text material related to political discussions and therefore created our own. We manually selected 46 debate articles from the Norwegian online newspapers \textit{NRK Ytring, Dagbladet, Aftenposten, VG and Bergens Tidene}. To each debate article there were attached a discussion thread where readers could express their opinions and feelings towards the content of the debate article. All the text from the debate articles and the subsequent discussions were collected using text scraping \cite{Hammer13}. The debate articles were related to religion and immigration and we wanted to classify the opinion towards all words with stem: \textit{islam, muslim, quran, allah, muhammed, imam and mosque}. These are topics that typically create a lot of active discussions and disagreements.

@@

Each sentence was manually annotated as to whether the commenter was positive, negative or neutral towards the keyword in the sentence. Each sentence was evaluated individually. The sentences were annotated based on real-world knowledge, e.g. a sentence like ``Muhammed is like Hitler'' would be annotated as a negative opinion towards Muhammed. Further, if a commenter presented a negative fact about the keyword, the sentence would be denoted as negative.

In order to assess inter-annotator agreement, a random sample of 65 sentences from the original text material was annotated by a second annotator. These sentences were not included in either the training or test set. For these sentences, the two annotators agreed on 58, which is an 89\% agreement, with a 95\% confidence interval equal to (79\%, 95\%) assuming that each sentence is independent. Since the sentences are drawn randomly from the population of all sentences this is a fair assumption.

Finally the material was divided into two parts where the first half of the debate articles with subsequent discussions make up the training set and the rest constitutes a held-out test set. In the development of the important dependency paths, only the training set was used. After the division, the training and test set consisted of a total of 382 and 308 sentences, respectively. Table \ref{tab:1} summarizes the opinions in the sentences
\begin{table}
  \caption{Manual annotation of training and test set.}
  \centering
  \begin{tabular}{lcccc}
             & Negative & Neutral & Positive \\\hline
    Training & 174 (46\%) & 162 (42\%) & 46 (12\%)\\
    Test     & 102 (33\%)& 182 (59\%) & 24 (8\%)
  \end{tabular}
  \label{tab:1}
\end{table}

@@

In this study we compared four different methods based on the general algorithm in \eqref{eq:1}.
\begin{itemize}
\item We use the \textbf{imp}-function presented in \eqref{eq:2}. We denote this method WD (word distance).
\item For this method and the two below we use the \textbf{imp}-function in \eqref{eq:3}. Further we set $\alpha = 0$ which means that we do not use the important dependency paths. We denote this method A0 ($\alpha = 0$).
\item We set $\alpha = 1$ and for all dependency paths we set $W_{\mathbf{dep}} = 2/3$. We denote this method CW (constant weights).
\item We set $\alpha = 1$ and for $W_{\mathbf{dep}}$ we use the weights presented in Table ??????. We denote this method OD (optimal use of dependency paths)
\end{itemize}
For each method we used the training set to manually tune the parameters $d_p, d_n, t_p$ and $t_n$ of the method. The parameters were tuned to optimize the number of correct classifications.

Table \ref{tab:2} shows the optimal parameter values of $d_p, d_n, t_p$ and $t_n$ tuned using the training set, and classification performance for the different methods on the test set using the parameter values tuned from the training set. The p-values are computed using the regression model presented in Section \ref{sec:sa}.
\begin{table}
  \caption{The second to the fifth column show the optimal values of the parameters of the model tuned using the training set. The sixth column show the number of correct classifications and the last column shows p-values testing whether the method performs better than WD.}
  \centering
  \begin{tabular}{lcccccl}
       & $d_p$ & $d_n$ & $t_p$ & $t_n$ & Correct   & p-val \\ \hline
    WD &    2  &   0  & 0.7  &  0.0  & 145 (47\%) & \\
    A0 &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.023\\
    CW &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.024\\
    OD &    2  &   0  & 2.0  &  0.3  & 162 (53\%) & 0.016
  \end{tabular}
  \label{tab:2}
\end{table}
We see that the sentiment shifter 'not' only have a positive effect on the classification performance when it is in front of the sentiment word. We see that using dependency distances (method A0) the classification results is significantly improved compared to using word distances in the sentence (method WD) (p-value = 0.023). Also classification based on important dependency paths (method OD) performs significantly better than WD. We also see that OD performs better than A0 (162 correct compared to 161), but this improvement is not statistically significant.

@@

Classifying opinions in political discussions is hard because of the frequent use of irony, sarcasm and citations. In this paper we have compared the use of word distance between keyword and sentiment word against metrics related to parsed sentence information. Our results show that using dependency tree distances or important dependency paths, improves the classification performance compared to using word distance.

Manually selecting important dependency paths for the aim of opinion mining is a hard task. A natural further step of our analysis is to expand the training and test material and use machine learning to see if there exists dependency paths that improve results compared to using dependency distance.
